### 2021-09-10 documents/config/own_abbreviations.tex
initial commit

\newGlossingAbbrev{inter}{interrogative}
\DeclareAcronym{tam}{short=TAM,long=Tense-Aspect-Mood,class=other}