### 2021-09-10 documents/floats/participles.tex

\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                               \gl{s_a_} &                        \gl{s_p_} \\
\midrule
\kaxui  &         \obj{t-ehurka-t͡ʃe} \qu{fallen} &       \obj{tɨ-jaʔ-so} \qu{burnt} \\
\arara  &         \obj{t-\emp{o-}ep-te} \qu{come} &      \obj{t-oregrum-te} \qu{sad} \\
\trio   &    \obj{tɨ-\emp{w-}əturu-e} \qu{talked} &       \obj{t-əpəə-se} \qu{tired} \\
\wayana &     \obj{tə-\emp{w-}epɨ-he} \qu{bathed} &    \obj{t-onopɨ-he} \qu{painted} \\
\apalai &  \obj{t-\emp{o-}erama-se} \qu{returned} &      \obj{t-anorɨ-se} \qu{dried} \\
\kalina &  \obj{tu-\emp{w-}oʔka-se} \qu{come out} &       \obj{t-okari-se} \qu{told} \\
\panare &  \obj{t-\emp{o-}tatɨhpə-se} \qu{wailed} &  \obj{tɨ-sɨrɨke-t͡ʃe} \qu{tired} \\
\bottomrule
\end{tabular}