### 2021-09-28 documents/floats/imperatives.tex
added imperative data, some optics

\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                         \gl{s_a_} &                             \gl{s_p_} \\
\midrule
\hixka  &          \obj{omoh-ko} \qu{come!} &  \obj{\emp{oj-}okajɨm-ko} \qu{go up!} \\
\arara  &  \obj{odotpot-ko} \qu{come back!} &      \obj{\emp{o-}alum-ko} \qu{jump!} \\
\trio   &          \obj{epɨ-kə} \qu{bathe!} &   \obj{\emp{ə-}eremina-kə} \qu{sing!} \\
\wayana &         \obj{əməm-kə} \qu{enter!} &    \obj{\emp{əw-}eremi-kə} \qu{sing!} \\
\apalai &           \obj{otuʔ-ko} \qu{eat!} &      \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} \\
\kalina &         \obj{oʔmaʔ-ko} \qu{stop!} &   \obj{\emp{aj-}awon-ko} \qu{get up!} \\
\panare &            \obj{ape-ʔ} \qu{flee!} &             \obj{ahpən-kə} \qu{jump!} \\
\bottomrule
\end{tabular}