### 2021-10-11 documents/floats/imperatives.tex

\begin{table}
\centering
\caption[Imperatives of \gl{s_a_} and \gl{s_p_} verbs]{Imperatives of \gl{s_a_} and \gl{s_p_} verbs \parencites[62]{koehn1986apalai}[Mopo/20]{koehns1994textos}[44, 89]{derbyshire1965textos}[5, 17]{mattei1994diccionario}[190]{hoff1968carib}[161]{alves2017arara}[323]{triomeira1999}[227]{wayanatavares2005}}
\label{tab:imperatives}
\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                         \gl{s_a_} &                             \gl{s_p_} \\
\midrule
\apalai &           \obj{otuʔ-ko} \qu{eat!} &      \obj{\emp{o-}nɨʔ-ko} \qu{sleep!} \\
\hixka  &          \obj{omoh-ko} \qu{come!} &  \obj{\emp{oj-}okajɨm-ko} \qu{go up!} \\
\panare &            \obj{ape-ʔ} \qu{flee!} &             \obj{ahpən-kə} \qu{jump!} \\
\kalina &         \obj{oʔmaʔ-ko} \qu{stop!} &   \obj{\emp{aj-}awon-ko} \qu{get up!} \\
\arara  &  \obj{odotpot-ko} \qu{come back!} &      \obj{\emp{o-}alum-ko} \qu{jump!} \\
\trio   &          \obj{epɨ-kə} \qu{bathe!} &   \obj{\emp{ə-}eremina-kə} \qu{sing!} \\
\wayana &         \obj{əməm-kə} \qu{enter!} &    \obj{\emp{əw-}eremi-kə} \qu{sing!} \\
\bottomrule
\end{tabular}
\end{table}