### 2021-10-11 documents/floats/nominalizations.tex

\begin{table}
\centering
\caption[Nominalizations of \gl{s_a_} and \gl{s_p_} verbs]{Nominalizations of \gl{s_a_} and \gl{s_p_} verbs \parencites[90]{koehn1986apalai}[ner2-003]{koehns1994textos}[390]{panarepayne2013}[23]{mattei1994diccionario}[135, 392]{hoff1968carib}[97]{alves2017arara}[49, 74]{schuring2018kaxuyana}[246]{triomeira1999}[130, 409]{wayanatavares2005}}
\label{tab:nominalizations}
\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &                                         \gl{s_a_} &                                             \gl{s_p_} \\
\midrule
\apalai &            \obj{j-epɨ-topo} \qu{my bathing place} &   \obj{j-enuru-topõ-pɨrɨ} \qu{the place of my birth} \\
\panare &     \obj{j-\emp{u-}t͡ʃireema-n} \qu{their eating} &  \obj{tj-arunkampətɨ-n} \qu{his hair standing on end} \\
\kalina &  \obj{a-\emp{w-}ekupi-rɨ} \qu{your taking a bath} &    \obj{aj-ereʔna-{\normalfont ∅}} \qu{your fainting} \\
\arara  &        \obj{\emp{w-}orik-tubo} \qu{dancing place} &              \obj{ereŋmɨ-tpo} \qu{killing instrument} \\
\kaxui  &       \obj{o-\emp{w-}ehurka-tpɨrɨ} \qu{your fall} &            \obj{o-onenmehɨ-tpɨrɨ} \qu{your waking up} \\
\trio   &   \obj{ji-\emp{w-}əturu-to} \qu{(for) my talking} &              \obj{j-emamina-to} \qu{(for) my playing} \\
\wayana &          \obj{ɨ-\emp{w-}əturu-topo} \qu{my story} &        \obj{j-ɨnɨkɨ-topo} \qu{my object for sleeping} \\
\bottomrule
\end{tabular}
\end{table}