### 2021-10-20 documents/floats/defecate.tex
cleaned up partial and normal alignment generation

\begin{table}
\centering
\caption[\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb]{\rc{weka} \qu{to defecate} as another class-switching \gl{s_p_} verb \parencites[418]{courtz2008carib}[455]{maquiritaricaceres2011}[44]{souza1993arara}[118]{alves2013verbo}[86, 206]{wayanatavares2005}[294]{triomeira1999}[96]{camargo2002lexico}[319; p.c., Spike Gildea]{mattei1994diccionario}}
\label{tab:defecate}
\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &         Form &                  Class \\
\midrule
\kaxui  &   \obj{weka} &              \gl{s_p_} \\
\arara  &  \obj{watke} &              \gl{s_p_} \\
\ikpeng &   \obj{atke} &              \gl{s_p_} \\
\trio   &   \obj{weka} &              \gl{s_a_} \\
\wayana &   \obj{uika} &  \gl{s_a_} / \gl{s_p_} \\
\apalai &   \obj{weka} &                      ? \\
\kalina &  \obj{uweka} &              \gl{s_p_} \\
\maqui  &   \obj{weka} &              \gl{s_p_} \\
\panare &   \obj{iʔka} &                      ? \\
\bottomrule
\end{tabular}
\end{table}