### 2021-10-25 documents/floats/say.tex

\begin{table}
\centering
\caption[Reflexes of \rc{ka(ti)} \qu{to say}]{Reflexes of \rc{ka(ti)} \qu{to say} \parencites[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209]{ikpengpacheco2001}[153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102; p.c., Spike Gildea]{mattei1994diccionario}}
\label{tab:say}
\begin{tabular}[t]{@{}llllll@{}}
\toprule
Language &          Form & \multicolumn{4}{l}{Alignment} \\
\midrule
\kaxui   &   \obj{ka[s]} &         k &  a &  s &    \\
\hixka   &   \obj{ka[h]} &         k &  a &  h &    \\
\waiwai  &   \obj{ka[s]} &         k &  a &  s &    \\
\arara   &      \obj{ke} &         k &  e &    &    \\
\ikpeng  &      \obj{ke} &         k &  e &    &    \\
\bakairi &      \obj{ke} &         k &  e &    &    \\
\trio    &      \obj{ka} &         k &  a &    &    \\
\akuriyo &      \obj{ka} &         k &  a &    &    \\
\carijo  &      \obj{ka} &         k &  a &    &    \\
\wayana  &   \obj{ka[i]} &         k &  a &    &  i \\
\apalai  &  \obj{ka[ʃi]} &         k &  a &  ʃ &  i \\
\kalina  &      \obj{ka} &         k &  a &    &    \\
\kapon   &      \obj{ka} &         k &  a &    &    \\
\pemon   &      \obj{ka} &         k &  a &    &    \\
\macushi &      \obj{ka} &         k &  a &    &    \\
\panare  &   \obj{ka[h]} &         k &  a &  h &    \\
\uxc     &      \obj{ki} &         k &  i &    &    \\
\bottomrule
\end{tabular}
\end{table}