### 2021-10-26 documents/floats/ptirreg.tex

\begin{table}
\centering
\caption[Regular \PTir \gl{s_a_} verbs]{Regular \PTir \gl{s_a_} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
\label{tab:ptirreg}
\begin{tabular}[t]{@{}lllllll@{}}
\mytoprule
{} & \multicolumn{3}{l}{\qu{to bathe (\gl{intr})}} & \multicolumn{3}{l}{\qu{to sleep}} \\
{} &                     \PTir &          \trio &        \akuriyo &             \PTir &              \trio &           \akuriyo \\
\midrule
\gl{1}   &             \rc{t͡ʃ-epɨ-} &   \obj{s-epɨ-} &  \obj{t͡ʃ-epɨ-} &    \rc{t-əənɨkɨ-} &    \obj{t-əənɨkɨ-} &    \obj{k-əənɨkɨ-} \\
\gl{2}   &               \rc{m-epɨ-} &   \obj{m-epɨ-} &    \obj{m-epɨ-} &    \rc{m-əənɨkɨ-} &    \obj{m-əənɨkɨ-} &    \obj{m-əənɨkɨ-} \\
\gl{1+2} &              \rc{ke-epɨ-} &  \obj{ke-epɨ-} &   \obj{ke-epɨ-} &  \rc{kɨt-əənɨkɨ-} &  \obj{kɨt-əənɨkɨ-} &  \obj{kəʔ-əənɨkɨ-} \\
\gl{3}   &               \rc{n-epɨ-} &   \obj{n-epɨ-} &    \obj{n-epɨ-} &    \rc{n-əənɨkɨ-} &    \obj{n-əənɨkɨ-} &    \obj{n-əənɨkɨ-} \\
\bottomrule
\end{tabular}
\end{table}