### 2021-10-26 documents/floats/pekreg.tex

\begin{table}
\centering
\caption[Regular Pekodian \gl{s_a_} verbs]{Regular Pekodian \gl{s_a_} verbs \parencites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
\label{tab:pekreg}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} & \bakairi \qu{to go up} &         \arara \qu{to dance} &            \ikpeng \qu{to run} \\
\midrule
\gl{1}   &           \obj{k-əku-} &               \obj{k-origu-} &                \obj{k-aranme-} \\
\gl{2}   &           \obj{m-əku-} &               \obj{m-origu-} &                \obj{m-aranme-} \\
\gl{1+2} &         \obj{kɨd-əku-} &             \obj{kud-origu-} &               \obj{kw-aranme-} \\
\gl{3}   &           \obj{n-əku-} &  \obj{{\normalfont ∅}-origu} &  \obj{{\normalfont ∅}-aranme-} \\
\bottomrule
\end{tabular}
\end{table}