### 2021-10-28 documents/floats/go.tex
further revision of verbs section

\begin{table}
\centering
\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[6; p.c., Spike Gildea]{franchetto2002kuikuro}}
\label{tab:go}
\begin{tabular}[t]{@{}lllllll@{}}
\mytoprule
Language & \multicolumn{6}{l}{Form} \\
\midrule
\kaxui    &     \obj{to[mo]} &    &  t &  o &  m &  o \\
\hixka    &      \obj{[ɨ]to} &  ɨ &  t &  o &    &    \\
\waiwai   &   \obj{[e]to[m]} &  e &  t &  o &  m &    \\
\arara    &        \obj{ɨdo} &  ɨ &  d &  o &    &    \\
\ikpeng   &       \obj{aran} &  a &  r &  a &  n &    \\
\ikpeng   &        \obj{ero} &  e &  r &  o &    &    \\
\bakairi  &      \obj{[ɨ]tə} &  ɨ &  t &  ə &    &    \\
\trio     &      \obj{tə[n]} &    &  t &  ə &  n &    \\
\akuriyo  &  \obj{[ə]tə[mɨ]} &  ə &  t &  ə &  m &  ɨ \\
\carijo   &       \obj{təmə} &    &  t &  ə &  m &  ə \\
\wayana   &   \obj{[ɨ]tə[m]} &  ɨ &  t &  ə &  m &    \\
\kalina   &         \obj{to} &    &  t &  o &    &    \\
\kalina   &       \obj{[ɨ]ʔ} &  ɨ &  ʔ &    &    &    \\
\maqui    &    \obj{ɨtə[mə]} &  ɨ &  t &  ə &  m &  ə \\
\ingariko &        \obj{ətə} &  ə &  t &  ə &    &    \\
\pemon    &      \obj{[e]tə} &  e &  t &  ə &    &    \\
\macushi  &      \obj{[a]tɨ} &  a &  t &  ɨ &    &    \\
\panare   &      \obj{tə[n]} &    &  t &  ə &  n &    \\
\yawarana &         \obj{tə} &    &  t &  ə &    &    \\
\mapoyo   &         \obj{tə} &    &  t &  ə &    &    \\
\uxc      &      \obj{[e]te} &  e &  t &  e &    &    \\
\yukpa    &         \obj{to} &    &  t &  o &    &    \\
\bottomrule
\end{tabular}
\end{table}