### 2021-11-07 documents/floats/hixintro.tex
tables for introduction

\begin{table}
\centering
\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510, 511, 513]{howard2001wrought}[27, 69, 70]{hixkaryanaderby1985}}
\label{tab:hixintro}
\begin{tabular}[t]{@{}lllll@{}}
\mytoprule
{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} &    \qu{to go} \\
\midrule
\gl{1}   &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &  \obj{k-atarʲeknohɨ-} &   \obj{ɨ-to-} \\
\gl{2}   &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{m-atarʲeknohɨ-} &  \obj{mɨ-to-} \\
\gl{1+2} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{t-atarʲeknohɨ-} &  \obj{tɨ-to-} \\
\gl{3}   &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &  \obj{n-atarʲeknohɨ-} &   \obj{n-to-} \\
\bottomrule
\end{tabular}
\end{table}