### 2021-11-12 documents/floats/triintro.tex

\begin{table}
\centering
\caption[Some \trio verbs]{Some \trio verbs \parencites[292, 294]{triomeira1999}[274]{triocarlin2004}}
\label{tab:triintro}
\begin{tabular}[t]{@{}llllll@{}}
\mytoprule
{} &      \qu{to sleep} & \qu{to see self} & \qu{to bathe (\gl{intr})} &     \qu{to yawn} &     \qu{to go} \\
\midrule
\gl{1}   &    \obj{t-əənɨkɨ-} &    \obj{t-əene-} &              \obj{s-epɨ-} &  \obj{s-entapo-} &  \obj{wɨ-tən-} \\
\gl{2}   &    \obj{m-əənɨkɨ-} &    \obj{m-əene-} &              \obj{m-epɨ-} &  \obj{m-entapo-} &  \obj{mɨ-tən-} \\
\gl{1+2} &  \obj{kɨt-əənɨkɨ-} &    \obj{k-əene-} &             \obj{ke-epɨ-} &  \obj{k-entapo-} &  \obj{kɨ-tən-} \\
\gl{3}   &    \obj{n-əənɨkɨ-} &    \obj{n-əene-} &              \obj{n-epɨ-} &  \obj{n-entapo-} &  \obj{nɨ-tən-} \\
\bottomrule
\end{tabular}
\end{table}