### 2021-11-23 documents/floats/ptirverbs.tex
started cleaning up workflow

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PTir \parencites[112, 113, 114, 115, 165]{meira1998proto}[292, 294, 339]{triomeira1999}}
\label{tab:ptirverbs}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &          \PTir &          \trio &                     \akuriyo \\
\midrule
\qu{be-1} &      \rc{w-a-} &     \obj{w-a-} &     \obj{{\normalfont ∅}-a-} \\
\qu{be-2} &    \rc{w-eʔi-} &    \obj{w-ei-} &   \obj{{\normalfont ?}-eʔi-} \\
\qu{say}  &    \rc{wɨ-ka-} &   \obj{wɨ-ka-} &                 \obj{wɨ-ka-} \\
\qu{go}   &  \rc{wɨ-təmɨ-} &  \obj{wɨ-tən-} &                \obj{ə-təmɨ-} \\
\qu{come} &  \rc{w-əʔepɨ-} &  \obj{w-əepɨ-} &  \obj{{\normalfont ∅}-eepɨ-} \\
\bottomrule
\end{tabular}
\end{table}