### 2021-11-23 documents/floats/pwaiverbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PWai \parencites[70, 197, 198]{hixkaryanaderby1985}[4]{hixkaryanaderby1979}[71, 85; p.c., Spike Gildea]{waiwaihawkins1998}}
\label{tab:pwaiverbs}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &         \PWai &        \hixka &         \waiwai \\
\midrule
\qu{be-1} &    \rc{w-ah-} &   \obj{w-ah-} &      \obj{w-a-} \\
\qu{be-2} &   \rc{w-eʃi-} &  \obj{w-eʃe-} &   \obj{w-eeʃi-} \\
\qu{say}  &   \rc{wɨ-ka-} &   \obj{ɨ-ka-} &    \obj{wɨ-ka-} \\
\qu{go}   &  \rc{wɨ-tom-} &   \obj{ɨ-to-} &  \obj{kɨw-tom-} \\
\bottomrule
\end{tabular}
\end{table}