### 2021-11-23 documents/floats/yukreg.tex

\begin{table}
\centering
\caption[Regular \yukpa verbs]{Regular \yukpa verbs \parencites[139]{meira2006syntactic}[72, 76]{largo2011yukpa}}
\label{tab:yukreg}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &  \qu{to fall} & \qu{to sleep} & \qu{to wash self} \\
\midrule
\gl{1} &  \obj{j-ata-} &  \obj{jɨ-nɨ-} &     \obj{j-otum-} \\
\gl{2} &  \obj{m-ata-} &  \obj{mɨ-nɨ-} &     \obj{m-otum-} \\
\gl{3} &  \obj{n-ata-} &  \obj{nɨ-nɨ-} &     \obj{n-otum-} \\
\bottomrule
\end{tabular}
\end{table}