### 2021-11-26 documents/floats/pxinw.tex

\begin{table}
\centering
\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[44, 70]{souza1993arara}[118]{alves2013verbo}[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[40]{campetela1997analise}}
\label{tab:pxinw}
\begin{tabular}[t]{@{}lll@{}}
\mytoprule
             Meaning &       \arara &     \ikpeng \\
\mymidrule
    \qu{to defecate} &  \obj{watke} &  \obj{atke} \\
       \qu{\gl{dat}} &   \obj{wɨna} &   \obj{ɨna} \\
            \qu{dog} & \obj{wokori} & \obj{akari} \\
\qu{capuchin monkey} &   \obj{tawe} &   \obj{tae} \\
       \qu{to sleep} &  \obj{wɨnkɨ} &  \obj{ɨnkɨ} \\
\mybottomrule
\end{tabular}
\end{table}