### 2021-11-27 documents/floats/yukreg.tex
incorporated new results into discussion

\begin{table}
\centering
\caption[Regular \yukpa verbs]{Regular \yukpa verbs \parencites[72, 76]{largo2011yukpa}[139]{meira2006syntactic}}
\label{tab:yukreg}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} & \qu{to sleep} & \qu{to wash self} &  \qu{to fall} \\
\mymidrule
\gl{1} &  \obj{jɨ-nɨ-} &     \obj{j-otum-} &  \obj{j-ata-} \\
\gl{2} &  \obj{mɨ-nɨ-} &     \obj{m-otum-} &  \obj{m-ata-} \\
\gl{3} &  \obj{nɨ-nɨ-} &     \obj{n-otum-} &  \obj{n-ata-} \\
\mybottomrule
\end{tabular}
\end{table}