### 2021-12-06 documents/sections/reasons.tex
rewriting section 4

\section{Explaining conservativeness: a network morphology approach}
\label{sec:motivations}
As stated in \cref{sec:verbsummary}, the fact that the same verbs retain conservative first person prefixes in six individual innovations is in need of explanation.
Perhaps the most well-known contribution regarding conservativeness, innovativeness, and (ir-){}re\-gu\-la\-ri\-ty in the lexicon is \posscite{bybee1985morphology} network model of morphology, which seems well-suited for the data at hand.
It aims \dbqu{to account for cross-linguistic, diachronic and acquisition patterns in complex morphological systems} \parencite[428]{bybee1995regular}.
It does so by modeling shared morphological properties such as inflectional patterns as emerging from connections of differing strength between lexemes.
A classic example is a network of \dbqu{strong} English verbs with \obj{strɪŋ}--\obj{strʌŋ} at the center and pairs like \obj{rɪŋ}--\obj{rʌŋ}, \obj{spɪn}--\obj{spʌn}, or \obj{stɪk}--\obj{stʌk} at its periphery.
This network is attracting new verbs in certain dialects, like \obj{sniːk}--\obj{snʌk} or \obj{brɪŋ}--\obj{brʌŋ} \parencite[129--130]{bybee1985morphology}.
These verbs are recruited based on the lexical connection they form with prototypical members of the group, and accordingly develop irregular or \dbqu{strong} past tense forms.

As possible bases of these connections between lexemes, \textcite[118]{bybee1985morphology} suggests the criteria of semantic, phonological, and morphological similarity; the English strong verbs are an example for a phonologically motivated network.
Another important factor in the model is frequency, since more frequent words have a higher lexical strength \parencite[119]{bybee1985morphology}.
This higher lexical strength diminishes the influence from other lexemes, meaning that high-frequency items are more likely to resist innovations.
For the Cariban inflectional patterns under investigation, the model would predict that a) semantically\slash{}phonologically\slash\hspace{0pt}morphologically similar verbs will be affected by person marker extensions, and b) high-frequency verbs will tend to resist these extensions and thus remain conservative.

When one considers the groups of verbs with innovative first person markers (those not in \cref{tab:overview}), one can easily perceive several factors which could serve as the thread connecting a lexical network.
Perhaps the most obvious similarity is that they all have a reflex of the detransitivizer \detrz, a hallmark of \gl{s_a_} verbs \pcref{sec:split}.
%This was already noted by \textcite{meira1998proto} for the group of \PTar verbs taking irregular first person \rc{w-}:
%\begin{quotebox}{\parencite[112]{meira1998proto}}
%	This category includes a small number of stems, among which ‘to go’, ‘to come’, ‘to say’, [...]\footnote{The original list includes \qu{to go down} and \qu{to defecate}. While these verbs are indeed underived \gl{s_a_} verbs in \trio, no irregular first person \rc{w-} can be reconstructed. As their inclusion in the list was most likely an error, we omit them here.} and the copula. These are basically the verbs that are not synchronically or diachronically detransitivized, yet belong to the A conjugation.
%\end{quotebox} % ‘to go down’, ‘to defecate’,
This also has a phonological consequence: all derived \gl{s_a_} verbs begin with reflexes of \rc{ə} or \rc{e}, meaning that networks with a phonological basis are also plausible.
A more trivial connection between these other verbs is that they are all \gl{s_a_} verbs, and thus share inflectional (morphological) patterns.
In the following analysis, the \gl{1}\gl{s_a_} (pre-innovation) prefix will serve as a proxy for fhese inflectional patterns.
This is made possible by the fact that \gl{s_a_} verbs all had a reflex of \rc{w-} at the time of innovation, making the first person prefix representative for the inflectional class.
The exception to this is \akuriyo, where innovative \gl{1}\gl{s_a_} \obj{k-} and idiosyncratic \obj{p-} on \qu{to go down} created inflectional subclasses of \gl{s_a_} verbs.
There are no semantic patterns that would be obvious, which is not unsurprising given the lack of semantic patterns in the split-\gl{s} system overall \pcref{sec:split}.
Thus, for each extension, there are three hypotheses as to what connected the members of the responsible network: a reflex of \gl{detrz}, their stem-initial phoneme, or a specific \gl{1}\gl{s_a_} prefix.

It is intuitively obvious that many of the conservative verbs in \cref{tab:overview} are high-frequency verbs, which would mean high lexical strength and conservativeness according to the network model.
A major obstacle to confirming this intuition is the lack of frequency counts for individual lexemes for any Cariban language.
The only statement in the literature is \posscite[75]{courtz2008carib} claim of \kalina underived \gl{s_a_} verbs being the most frequent ones: \dbqu{It is difficult […] to imagine an intransitive or transitive origin for some of the most frequent middle verbs}.
This claim is supported neither by frequency counts nor accompanied by a list of verbs, but the unmentioned verbs likely correspond to the five first columns of \cref{tab:overview}, all underived \gl{s_a_} verbs in \kalina.
To ameliorate the dearth of data, a count of \gl{s_a_} verbs in three glossed texts from \textcite{koehns1994textos} will serve as a second source of frequency information, the results of which are shown in \cref{tab:apalaicounts}.
The \apalai data agree with the above interpretation of \posscite{courtz2008carib} claim; defining \dbqu{high frequency} as having an above average count produces the exact same five verbs.
While it is not at all certain that the interpretation of \posscite{courtz2008carib} claim and the small \apalai sample are representative of discourse patterns in the Cariban (proto-)languages under discussion, their use as a tool for identifying high-frequency verbs is necessitated by the absence of alternatives.

\input{floats/apalaicounts.tex}

%\textcite[153]{alves2017arara}: \dbqu{Observe que todos os 12 verbos indexados pelo morfema \obj{k-} [\gl{s_a_}] apresentam a raiz iniciada pela vogal /o/}.

Thus, in addition to the three hypotheses for possible network factors, each can be combined with frequency; high-frequency verbs are predicted to not undergo innovation, even though the factor in question would put them in the same lexical network as regular \gl{s_a_} verbs.
This leaves each extension with six possible explanations for which verbs are affected and which are not.
To test the hypotheses, for each explanation the expected behavior of each verb was predicted, illustrated in \cref{tab:ptir-predictions} for \PTir.
For example, \rc{eʔi} \qu{to be} is expected to participate in innovations spreading in a phonologically defined network, being \rc{e}-initial, as well as in an inflectionally defined one, since it shared \rc{w-} with other \gl{s_a_} verbs.
However, it would not have belonged to a network defined by the presence of a detransitivizer.
If frequency is taken into account, it is expected to remain conservative regardless of the nature of the network.
These predictions were then checked against the data in \cref{tab:overview}, to see how many potentially conservative verb each explanation predicted correctly.
This yielded a score of what proportion of potentially conservative verbs had their behavior accurately predicted, illustrated for \PTir in \cref{tab:ptir-evaluations} and summed up in \cref{tab:resultsoverview}.

\input{floats/ptir-predictions.tex}
\input{floats/ptir-evaluations.tex}

\input{floats/resultsoverview.tex}

It is important to understand that the scores in \cref{tab:resultsoverview} only refer to the group of seven verbs in \cref{tab:overview}, i.e. those that are attested as resisting at least one extension.
For each extension, there were also many run-of-the-mill \gl{s_a_} verbs, all taking on the new person marker, except for the \akuriyo \obj{e}-initial verbs.%
\footnote{While there are a few \gl{s_a_} verbs not transparently derived from transitive verbs \parencites[252]{triomeira1999}[222]{meira2000split}[30]{gildea2007greenberg}, which are not featured in \cref{tab:overview}, these are mostly \rc{ə}-initial and were likely productively derived at some point.
	The verbs to which this does not apply, like \trio \obj{wa} \qu{to dance} \parencites[252]{triomeira1999}, are all instances of \gl{s_p_} verbs switching classes.
	Since none of them is attested as being an \gl{s_a_} verb at the time of a person marker extension, they are not relevant for this study.}
To illustrate, if one adds 1'000 regular \gl{s_a_} verbs per language to the data -- a conservative estimate based on \posscite{courtz2008carib} \kalina dictionary -- all six explanations consistently predict the behavior of 99.99+\% verbs correctly.
However, the available data simply does not allow such large-scale tests for Cariban languages, so the present investigation is restricted to the edge cases.

The extent of the extensions in both \PWai and \PPek is fully predicted by the presence or absence of a detransitivizer.
In both cases, only the underived%
\footnote{Note that for \PPek, the idiosyncratic evolution of \rc{e-pɨ} \qu{to bathe (\gl{intr})} to \rc{ipɨ}  made the verb morphologically opaque.}
\gl{s_a_} verbs were not affected, all other \gl{s_a_} verbs taking \rc{k-}.
Not shown in \cref{tab:resultsoverview} are subsequent evolutions in the Pekodian daughter languages, which can largely be argued to also be due to the detransitivizer:
First, both \ikpeng and \bakairi regularized the paradigm to use forms with detransitivizer for first person \pcref{sec:pekodian}, and both introduced \obj{k-}.%
\footnote{If one instead assumes that first person \rc{w-ebɨ-} and \rc{k-əd-ebɨ-} already co-existed in \PPek, the clear correlation between \rc{k-} and the detransitivizer remains.}
Second, the subsequent introduction of \obj{k-} to \ikpeng \obj{aran} \qu{to go} (< \rc{ɨtən}) suggests a potential reanalysis of \obj{ar} as a detransitivizer.

Three extensions are fully predicted by phonological criteria, those in \akuriyo, \carijo, and \yukpa.
As already discussed, \akuriyo \obj{k-} only appears on \obj{ə}-initial verbs \pcref{sec:akuriyo}.
In \carijo, the extension of \obj{j-} affected all \obj{e}- and \obj{ə}-initial verbs, including \obj{eh} \qu{to come} or \obj{et͡ʃi} \qu{to be}, which do not have a detransitivizing prefix.
Only \obj{ka} \qu{say}, \obj{təmə} \qu{go}, and \obj{a} \qu{be-1} did not take on \obj{j-}.
Similarly, the extension of \yukpa \obj{j-} can succinctly be characterized as affecting all vowel-initial verbs; the only attested conservative verb is C-initial \obj{to} \qu{to go}.

Inflectional morphology as a network basis only played a potential role in the case of \akuriyo.
However, this explanation only works if \obj{t-} and \obj{t͡ʃ-}, which were of course also phonologically conditioned, are analyzed as distinct morphemes.

When additionally considering putative conservatory frequency effects, prediction accuracy was improved in 8 cases, stagnated in 7 cases, and worsened in 3 cases.
The three cases where the tentative model of verb frequency arrives at incorrect predictions are found in \carijo and \yukpa, the only languages to feature innovative markers on the reflexes of \rc{eti} \qu{be-2} or \rc{a[p]} \qu{be-1}.
Including frequency in the model led to overall improvements, up to 100\% prediction accuracy for all three potential factors in \PTir, as well as for the phonological criterion in \PPek and the inflection criterion in \PWai.

Overall, one must conclude that the patterns of most extensions are correctly predicted not by a single explanation, but rather by 3 to 4 different ones.
The exceptions are \carijo and \yukpa, where a lexical network with a phonolgical basis emerges as an unambiguous winner, while frequency-based explanations fare much worse.
%It may be worth noting that the three extensions which we do not reconstruct to a proto-language (ti.e., which are more recent) are best explained by phonological conditioning factors.
For the other extensions, the network model gives no unambiguous answer to the question of what combination of factors caused innovative first person markers to spread the way they did.
This in turn is due to the fact that three of the factors used in the model -- detransitivizer, phonology, frequency -- largely converge in their predictions:
The most frequent verbs are at the same time those without a detransitivizer, and therefore mostly of a different phonological shape than regular \gl{s_a_} verbs.