### 2021-12-16 documents/floats/kaxprog.tex

\begin{table}
\centering
\caption[\kaxui \gl{s_a_} verbs in the Progressive]{\kaxui \gl{s_a_} verbs in the Progressive (Spike Gildea, p.c.)}
\label{tab:kaxprog}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &                   \qu{to come} &                   \qu{to dream} &                    \qu{to go} \\
\mymidrule
\gl{1}   &  \obj{{\normalfont ∅}-w-oohɨ-} &  \obj{{\normalfont ∅}-w-osone-} &  \obj{{\normalfont ∅}-wɨ-to-} \\
\gl{2}   &                 \obj{o-w-ohɨ-} &                \obj{o-w-osone-} &                 \obj{o-w-to-} \\
\gl{1+2} &                \obj{ku-w-ohɨ-} &               \obj{ku-w-osone-} &                \obj{kɨ-w-to-} \\
\gl{3}   &    \obj{{\normalfont ∅}-johɨ-} &    \obj{{\normalfont ∅}-osone-} &                   \obj{i-to-} \\
\mybottomrule
\end{tabular}
\end{table}