### 2021-12-17 documents/floats/comepara.tex

\begin{table}
\centering
\caption[\rc{(ət-)epɨ} \qu{to come} in paradigms]{\rc{(ət-)epɨ} \qu{to come} in paradigms \parencites[113, 150, 153, 156]{alves2017arara}[294; p.c., Spike Gildea]{triomeira1999}}
\label{tab:comepara}
\begin{tabular}[t]{@{}llll@{}}
\mytoprule
{} &          \kaxui &          \arara &          \trio \\
\mymidrule
\gl{1}   &   \obj{w-oohɨ-} &    \obj{w-ebɨ-} &  \obj{w-əepɨ-} \\
\gl{2}   &   \obj{m-oohɨ-} &  \obj{m-odebɨ-} &  \obj{mən-epɨ} \\
\gl{1+2} &  \obj{kɨs-ohɨ-} &  \obj{kud-ebɨ-} &   \obj{ke-epɨ} \\
\gl{3}   &    \obj{n-ehɨ-} &    \obj{t-ebɨ-} &    \obj{n-epɨ} \\
\mybottomrule
\end{tabular}
\end{table}