### 2021-12-17 documents/floats/hixintro.tex

\begin{table}
\centering
\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510-511, 513, 520]{howard2001wrought}[197-198]{hixkaryanaderby1985}}
\label{tab:hixintro}
\begin{tabular}[t]{@{}llllll@{}}
\mytoprule
{} &     \qu{to fall} &  \qu{to be afraid} &          \qu{to walk} & \qu{to cut self} &    \qu{to be} \\
\mymidrule
\gl{1}   &  \obj{k-ehurka-} &  \obj{k-oserʲehɨ-} &  \obj{k-atarʲeknohɨ-} &   \obj{k-atama-} &  \obj{w-eʃe-} \\
\gl{2}   &  \obj{m-ehurka-} &  \obj{m-oserʲehɨ-} &  \obj{m-atarʲeknohɨ-} &   \obj{m-atama-} &  \obj{m-eʃe-} \\
\gl{1+2} &  \obj{t-ehurka-} &  \obj{t-oserʲehɨ-} &  \obj{t-atarʲeknohɨ-} &   \obj{t-atama-} &  \obj{t-eʃe-} \\
\gl{3}   &  \obj{ɲ-ehurka-} &  \obj{n-oserʲehɨ-} &  \obj{n-atarʲeknohɨ-} &   \obj{n-atama-} &  \obj{n-eʃe-} \\
\mybottomrule
\end{tabular}
\end{table}