### 2021-12-17 documents/floats/ppekverbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{s_a_} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[76, 130-131, 374, 397]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}}
\label{tab:ppekverbs}
\begin{tabular}[t]{@{}lllll@{}}
\mytoprule
{} &          \PPek &          \arara &                       \ikpeng &        \bakairi \\
\mymidrule
\qu{say}     &    \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &     \obj{u-ge-} \\
\qu{go}      &   \rc{w-ɨtən-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &     \obj{u-tə-} \\
\qu{be-1}    &     \rc{w-ap-} &     \obj{w-ap-} &                             – &      \obj{w-a-} \\
\qu{be-2}    &  \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &      \obj{w-i-} \\
\qu{come}    &    \rc{w-epɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &   \obj{k-əewɨ-} \\
\qu{go down} &   \rc{w-ɨptə-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ?}-iptoŋ-} &  \obj{k-ɨtəgɨ-} \\
\qu{bathe}   &    \rc{w-ipɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &      \obj{w-i-} \\
\mybottomrule
\end{tabular}
\end{table}