### 2021-12-25 documents/config/cadernos.tex
journal-specific formatting

\documentclass[a4paper, 11pt, twoside]{article}

%LAYOUT
\usepackage[left=2.4cm,right=2.4cm, top=2.5cm, bottom=2.4cm]{geometry}

%FONTS
\usepackage{fontspec}
\setmainfont{Linux Libertine}

%GLOSSES
\usepackage{glossingtool}
\lingset{aboveglftskip=-0.3ex, everygla=\itshape, belowglpreambleskip=-0.5ex, everyglpreamble=\itshape}
\input{config/own_abbreviations.tex}

%FIRST MENTIONS
\newcommand{\firstmention}[1]{\textsc{#1}}