### 2021-12-25 documents/config/linguistic_discovery.tex

\documentclass[letterpaper, 12pt]{article}

%LAYOUT
\usepackage[left=2.54cm, right=2.54cm, top=2.54cm, bottom=2.54cm]{geometry}

%SPACING
%\usepackage{setspace}
%\doublespacing

%GLOSSES
\usepackage{glossingtool}
\definelingstyle{IJAL}{labelwidth=2em,labelanchor=numleft, labeloffset=0pt,labelformat=(A),everylabel=\actualexno, textanchor=normal,textoffset=1em,preambleanchor=text, preambleoffset=0pt,avoidnumlabelclash, appendtopexarg={samplelabel=(\actualexno a)}}
\lingset{aboveglftskip=-0.3ex, everygla=\itshape, belowglpreambleskip=-0.5ex, everyglpreamble=\itshape, lingstyle=IJAL}
\input{config/own_abbreviations.tex}

%FONTS
\usepackage{fontspec}
\setmainfont{Linux Libertine}
\newfontfamily\objfont{Linux Libertine} %Doulos SIL does not have italics...
\renewcommand{\obj}[1]{\textit{\objfont#1}}
\renewcommand{\rc}[1]{*\textit{\objfont#1}}
\renewcommand{\ort}[1]{$\langle$\objfont#1$\rangle$}

\lingset{everygla=\itshape\objfont, everyglpreamble=\itshape\objfont, everyglb=\objfont}
\usepackage{newunicodechar}
\newunicodechar{∅}{{\objfont∅}}

%HEADERS AND FOOTERS
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead[L,C,R]{}
\renewcommand{\headrulewidth}{0pt}
%\fancyhead[R]{}
%\fancyfoot[L]{}
\fancyfoot[C]{\thepage}
%\setlength{\headheight}{14.5pt}

%SECTION HEADERS
\usepackage{titlesec}
\titleformat{\section}
{\normalfont\fontsize{14}{14}\bfseries}{\thesection.}{1ex}{}

\titleformat{\subsection}
{\normalfont\fontsize{12}{12}\bfseries}{\thesubsection.}{1ex}{}

%FIRST MENTIONS
\newcommand{\firstmention}[1]{\textit{#1}}

%TITLE AND ABSTRACT
\date{}
\usepackage{etoolbox}
\AtBeginEnvironment{abstract}{\itshape}

\usepackage{placeins}

%FLOATS AND FOOTNOTES
\usepackage[bottom]{footmisc}

% LINENUMBERS
%\usepackage{lineno}
\newcommand{\linenumbers}{}