### 2023-01-16 documents/config/own_abbreviations.tex
latex compilation working again

\newGlossingAbbrev{inter}{interrogative}
\DeclareAcronym{tam}{short=TAM,long=Tense-Aspect-Mood}