### 2023-01-16 documents/config/cadernos.tex
font, 'alignment'

\documentclass[a4paper, 11pt, twoside]{article}

%LAYOUT
\usepackage[left=2.4cm,right=2.4cm, top=2.5cm, bottom=2.4cm]{geometry}

%FONTS
\usepackage{fontspec}
\setmainfont{Brill}

%GLOSSES
\usepackage[abbrevs=all]{expex-acro}
\usepackage{multicol}
\lingset{aboveglftskip=-0.3ex, everygla=\itshape, belowglpreambleskip=-0.5ex, everyglpreamble=\itshape}
\input{config/own_abbreviations.tex}

%FIRST MENTIONS
\newcommand{\firstmention}[1]{\textsc{#1}}

%HEADERS AND FOOTERS
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead[]{\thepage}
\chead[Conservative first person forms in Cariban]{Cadernos de Etnolingüística}
\rhead[\thepage]{}
\setlength{\headheight}{15pt}

\lfoot{}
\cfoot{}
\rfoot{}

%LINE NUMBERS
\usepackage{lineno}
%\newcommand{\linenumbers}{}

%SPACING
%\usepackage{setspace}
%\onehalfspacing

%FLOATS AND FOOTNOTES
\usepackage[bottom]{footmisc}