### 2023-08-23 documents/config/cariban_language_names.tex
more table fixes

\usepackage{xspace}
\providecommand{\wbakairi}{Western Bakairi\xspace}
\DeclareAcronym{wbak_lg}{short=wbak,long=Western Bakairi,short-format=\scshape}
\providecommand{\ebakairi}{Eastern Bakairi\xspace}
\DeclareAcronym{ebak_lg}{short=ebak,long=Eastern Bakairi,short-format=\scshape}
\providecommand{\ingariko}{Ingarikó\xspace}
\DeclareAcronym{ing_lg}{short=ing,long=Ingarikó,short-format=\scshape}
\providecommand{\akawaio}{Akawaio\xspace}
\DeclareAcronym{aka_lg}{short=aka,long=Akawaio,short-format=\scshape}
\providecommand{\patamona}{Patamona\xspace}
\DeclareAcronym{pat_lg}{short=pat,long=Patamona,short-format=\scshape}
\providecommand{\tabajari}{Tabajari\xspace}
\DeclareAcronym{tab_lg}{short=tab,long=Tabajari,short-format=\scshape}
\providecommand{\gkar}{Guianese Carib\xspace}
\DeclareAcronym{gkar_lg}{short=gkar,long=Guianese Carib,short-format=\scshape}
\providecommand{\aretyry}{Aretyry\xspace}
\DeclareAcronym{aty_lg}{short=aty,long=Aretyry,short-format=\scshape}
\providecommand{\tyrewuju}{Tyrewuju\xspace}
\DeclareAcronym{tyr_lg}{short=tyr,long=Tyrewuju,short-format=\scshape}
\providecommand{\sikiana}{Sikiana\xspace}
\DeclareAcronym{siku_lg}{short=siku,long=Sikiana,short-format=\scshape}
\providecommand{\dekwana}{De’kwana\xspace}
\DeclareAcronym{dek_lg}{short=dek,long=De’kwana,short-format=\scshape}
\providecommand{\dhekwana}{Dhe’kwana\xspace}
\DeclareAcronym{dhe_lg}{short=dhe,long=Dhe’kwana,short-format=\scshape}
\providecommand{\kamara}{Kamarakoto\xspace}
\DeclareAcronym{kam_lg}{short=kam,long=Kamarakoto,short-format=\scshape}
\providecommand{\tauli}{Taulipang\xspace}
\DeclareAcronym{tau_lg}{short=tau,long=Taulipang,short-format=\scshape}
\providecommand{\arekuna}{Arekuna\xspace}
\DeclareAcronym{are_lg}{short=are,long=Arekuna,short-format=\scshape}
\providecommand{\nahukwa}{Nahukwa\xspace}
\DeclareAcronym{nah_lg}{short=nah,long=Nahukwa,short-format=\scshape}
\providecommand{\kuikuro}{Kuikuro\xspace}
\DeclareAcronym{kui_lg}{short=kui,long=Kuikuro,short-format=\scshape}
\providecommand{\matipu}{Matipu\xspace}
\DeclareAcronym{mat_lg}{short=mat,long=Matipu,short-format=\scshape}
\providecommand{\kalapalo}{Kalapalo\xspace}
\DeclareAcronym{kal_lg}{short=kal,long=Kalapalo,short-format=\scshape}
\providecommand{\yukpa}{Yukpa\xspace}
\DeclareAcronym{yuk_lg}{short=yuk,long=Yukpa,short-format=\scshape}
\providecommand{\japreria}{Japreria\xspace}
\DeclareAcronym{jap_lg}{short=jap,long=Japreria,short-format=\scshape}
\providecommand{\carijo}{Karijona\xspace}
\DeclareAcronym{car_lg}{short=car,long=Karijona,short-format=\scshape}
\providecommand{\mapoyo}{Mapoyo\xspace}
\DeclareAcronym{map_lg}{short=map,long=Mapoyo,short-format=\scshape}
\providecommand{\yawarana}{Yawarana\xspace}
\DeclareAcronym{yab_lg}{short=yab,long=Yawarana,short-format=\scshape}
\providecommand{\panare}{Panare\xspace}
\DeclareAcronym{pan_lg}{short=pan,long=Panare,short-format=\scshape}
\providecommand{\maqui}{Ye'kwana\xspace}
\DeclareAcronym{mak_lg}{short=mak,long=Ye'kwana,short-format=\scshape}
\providecommand{\pemon}{Pemón\xspace}
\DeclareAcronym{pem_lg}{short=pem,long=Pemón,short-format=\scshape}
\providecommand{\kapon}{Kapong\xspace}
\DeclareAcronym{kap_lg}{short=kap,long=Kapong,short-format=\scshape}
\providecommand{\waimiri}{Waimiri-Atroari\xspace}
\DeclareAcronym{wmr_lg}{short=wmr,long=Waimiri-Atroari,short-format=\scshape}
\providecommand{\macushi}{Macushi\xspace}
\DeclareAcronym{mac_lg}{short=mac,long=Macushi,short-format=\scshape}
\providecommand{\waiwai}{Waiwai\xspace}
\DeclareAcronym{wai_lg}{short=wai,long=Waiwai,short-format=\scshape}
\providecommand{\hixka}{Hixkaryána\xspace}
\DeclareAcronym{hix_lg}{short=hix,long=Hixkaryána,short-format=\scshape}
\providecommand{\kaxui}{Werikyana\xspace}
\DeclareAcronym{kax_lg}{short=kax,long=Werikyana,short-format=\scshape}
\providecommand{\trio}{Tiriyó\xspace}
\DeclareAcronym{tri_lg}{short=tri,long=Tiriyó,short-format=\scshape}
\providecommand{\akuriyo}{Akuriyó\xspace}
\DeclareAcronym{aku_lg}{short=aku,long=Akuriyó,short-format=\scshape}
\providecommand{\apalai}{Apalaí\xspace}
\DeclareAcronym{apa_lg}{short=apa,long=Apalaí,short-format=\scshape}
\providecommand{\wayana}{Wayana\xspace}
\DeclareAcronym{way_lg}{short=way,long=Wayana,short-format=\scshape}
\providecommand{\ikpeng}{Ikpeng\xspace}
\DeclareAcronym{ikp_lg}{short=ikp,long=Ikpeng,short-format=\scshape}
\providecommand{\arara}{Arara\xspace}
\DeclareAcronym{ara_lg}{short=ara,long=Arara,short-format=\scshape}
\providecommand{\kalina}{Kari'ña\xspace}
\DeclareAcronym{kar_lg}{short=kar,long=Kari'ña,short-format=\scshape}
\providecommand{\uxc}{Upper Xingu Carib\xspace}
\DeclareAcronym{uxc_lg}{short=uxc,long=Upper Xingu Carib,short-format=\scshape}
\providecommand{\bakairi}{Bakairi\xspace}
\DeclareAcronym{bak_lg}{short=bak,long=Bakairi,short-format=\scshape}
\providecommand{\hiana}{†Hianakoto-Umawa\xspace}
\DeclareAcronym{hian_lg}{short=hian,long=†Hianakoto-Umawa,short-format=\scshape}
\providecommand{\chayma}{†Chayma\xspace}
\DeclareAcronym{cha_lg}{short=cha,long=†Chayma,short-format=\scshape}
\providecommand{\kumanagoto}{†Kumanagoto\xspace}
\DeclareAcronym{cgt_lg}{short=cgt,long=†Kumanagoto,short-format=\scshape}
\providecommand{\arina}{†Arinakoto\xspace}
\DeclareAcronym{arin_lg}{short=arin,long=†Arinakoto,short-format=\scshape}
\providecommand{\saluma}{†Salumá\xspace}
\DeclareAcronym{salu_lg}{short=salu,long=†Salumá,short-format=\scshape}
\providecommand{\urukue}{†Urukuena\xspace}
\DeclareAcronym{uruk_lg}{short=uruk,long=†Urukuena,short-format=\scshape}
\providecommand{\wayakule}{†Wayakulé\xspace}
\DeclareAcronym{waya_lg}{short=waya,long=†Wayakulé,short-format=\scshape}
\providecommand{\pianakoto}{†Pianakoto\xspace}
\DeclareAcronym{pian_lg}{short=pian,long=†Pianakoto,short-format=\scshape}
\providecommand{\kumayena}{†Kumayena\xspace}
\DeclareAcronym{kuma_lg}{short=kuma,long=†Kumayena,short-format=\scshape}
\providecommand{\parukoto}{†Parukoto\xspace}
\DeclareAcronym{paru_lg}{short=paru,long=†Parukoto,short-format=\scshape}
\providecommand{\yawaperi}{†Jawaperi\xspace}
\DeclareAcronym{jawa_lg}{short=jawa,long=†Jawaperi,short-format=\scshape}
\providecommand{\macoita}{†Macoita\xspace}
\DeclareAcronym{mcta_lg}{short=mcta,long=†Macoita,short-format=\scshape}
\providecommand{\macoa}{†Macoa\xspace}
\DeclareAcronym{maco_lg}{short=maco,long=†Macoa,short-format=\scshape}
\providecommand{\chake}{†Chake\xspace}
\DeclareAcronym{chak_lg}{short=chak,long=†Chake,short-format=\scshape}
\providecommand{\koyaima}{†Koyaima\xspace}
\DeclareAcronym{koya_lg}{short=koya,long=†Koyaima,short-format=\scshape}
\providecommand{\opon}{†Opón-Carare\xspace}
\DeclareAcronym{opon_lg}{short=opon,long=†Opón-Carare,short-format=\scshape}
\providecommand{\guake}{†Guake\xspace}
\DeclareAcronym{guak_lg}{short=guak,long=†Guake,short-format=\scshape}
\providecommand{\pemono}{†Pémono\xspace}
\DeclareAcronym{pno_lg}{short=pno,long=†Pémono,short-format=\scshape}
\providecommand{\piritu}{†Píritu\xspace}
\DeclareAcronym{piri_lg}{short=piri,long=†Píritu,short-format=\scshape}
\providecommand{\tamanaku}{†Tamanaku\xspace}
\DeclareAcronym{tam_lg}{short=tam,long=†Tamanaku,short-format=\scshape}
\providecommand{\palenque}{†Palenque\xspace}
\DeclareAcronym{pale_lg}{short=pale,long=†Palenque,short-format=\scshape}
\providecommand{\avarikoto}{†Avarikoto\xspace}
\DeclareAcronym{avar_lg}{short=avar,long=†Avarikoto,short-format=\scshape}
\providecommand{\kumana}{†Kumaná\xspace}
\DeclareAcronym{cum_lg}{short=cum,long=†Kumaná,short-format=\scshape}
\providecommand{\palmela}{†Palmela\xspace}
\DeclareAcronym{palm_lg}{short=palm,long=†Palmela,short-format=\scshape}
\providecommand{\pauxiana}{†Pauxiana\xspace}
\DeclareAcronym{pxan_lg}{short=pxan,long=†Pauxiana,short-format=\scshape}
\providecommand{\tiveri}{†Tiverighotto\xspace}
\DeclareAcronym{tive_lg}{short=tive,long=†Tiverighotto,short-format=\scshape}
\providecommand{\purukoto}{†Purukoto\xspace}
\DeclareAcronym{puru_lg}{short=puru,long=†Purukoto,short-format=\scshape}
\providecommand{\sapara}{†Sapara\xspace}
\DeclareAcronym{sapa_lg}{short=sapa,long=†Sapara,short-format=\scshape}
\providecommand{\yao}{†Yao\xspace}
\DeclareAcronym{yao_lg}{short=yao,long=†Yao,short-format=\scshape}
\providecommand{\wayu}{†Wayumara\xspace}
\DeclareAcronym{wayu_lg}{short=wayu,long=†Wayumara,short-format=\scshape}
\providecommand{\juma}{†Juma\xspace}
\DeclareAcronym{juma_lg}{short=juma,long=†Juma,short-format=\scshape}
\providecommand{\paravi}{†Paravilhana\xspace}
\DeclareAcronym{para_lg}{short=para,long=†Paravilhana,short-format=\scshape}
\providecommand{\crichana}{†Crichana\xspace}
\DeclareAcronym{cric_lg}{short=cric,long=†Crichana,short-format=\scshape}
\providecommand{\bonari}{†Bonari\xspace}
\DeclareAcronym{bona_lg}{short=bona,long=†Bonari,short-format=\scshape}
\providecommand{\pauxi}{†Pauxi\xspace}
\DeclareAcronym{paux_lg}{short=paux,long=†Pauxi,short-format=\scshape}
\providecommand{\yaruma}{†Yarumá\xspace}
\DeclareAcronym{yaru_lg}{short=yaru,long=†Yarumá,short-format=\scshape}
\providecommand{\arakaju}{†Arakaju\xspace}
\DeclareAcronym{arak_lg}{short=arak,long=†Arakaju,short-format=\scshape}
\providecommand{\apiaka}{†Apiaka-Tocantins\xspace}
\DeclareAcronym{apia_lg}{short=apia,long=†Apiaka-Tocantins,short-format=\scshape}
\providecommand{\piment}{†Pimenteira\xspace}
\DeclareAcronym{pime_lg}{short=pime,long=†Pimenteira,short-format=\scshape}
\providecommand{\PC}{Proto-Cariban\xspace}
\DeclareAcronym{PC_lg}{short=PC,long=Proto-Cariban,short-format=\scshape}
\providecommand{\PPar}{Proto-Parukotoan\xspace}
\DeclareAcronym{PPar_lg}{short=PPar,long=Proto-Parukotoan,short-format=\scshape}
\providecommand{\PWai}{Proto-Waiwaian\xspace}
\DeclareAcronym{PWai_lg}{short=PWai,long=Proto-Waiwaian,short-format=\scshape}
\providecommand{\PPek}{Proto-Pekodian\xspace}
\DeclareAcronym{PPek_lg}{short=PPek,long=Proto-Pekodian,short-format=\scshape}
\providecommand{\PXin}{Proto-Xinguan\xspace}
\DeclareAcronym{PXin_lg}{short=PXin,long=Proto-Xinguan,short-format=\scshape}
\providecommand{\PPP}{Proto-Pemong-Panare\xspace}
\DeclareAcronym{PPP_lg}{short=PPP,long=Proto-Pemong-Panare,short-format=\scshape}
\providecommand{\PPem}{Proto-Pemongan\xspace}
\DeclareAcronym{PPem_lg}{short=PPem,long=Proto-Pemongan,short-format=\scshape}
\providecommand{\PTar}{Proto-Taranoan\xspace}
\DeclareAcronym{PTar_lg}{short=PTar,long=Proto-Taranoan,short-format=\scshape}
\providecommand{\PTir}{Proto-Tiriyoan\xspace}
\DeclareAcronym{PTir_lg}{short=PTir,long=Proto-Tiriyoan,short-format=\scshape}
\providecommand{\PMan}{Proto-Manapiarean\xspace}
\DeclareAcronym{PMan_lg}{short=PMan,long=Proto-Manapiarean,short-format=\scshape}
\providecommand{\PVen}{Proto-Venezuelan\xspace}
\DeclareAcronym{PVen_lg}{short=PVen,long=Proto-Venezuelan,short-format=\scshape}
\providecommand{\PCum}{Proto-Cumana\xspace}
\DeclareAcronym{PCum_lg}{short=PCum,long=Proto-Cumana,short-format=\scshape}