### 2023-08-23 documents/floats/aku1sa.tex

\begin{table}
\centering
\caption[Regular \akuriyo \gl{1}\gl{s_a_} markers]{Regular \akuriyo \gl{1}\gl{s_a_} markers \parencites[77, 79, 82, 84-86]{gildea1994akuriyo}}
\label{tab:aku1sa}
\begin{tabular}[t]{@{}ll@{}}
\mytoprule
      first person \obj{k-} &               first person \obj{t͡ʃ-} \\
\mymidrule
 \obj{əempa-} \qu{to learn} &            \obj{entapo-} \qu{to yawn} \\
\obj{əiwa-} \qu{to tremble} &      \obj{ekɨrɨka-} \qu{to stay back} \\
\obj{ətajiŋka-} \qu{to run} &       \obj{etonema-} \qu{to lie down} \\
\obj{əət͡ʃena-} \qu{to cry} &          \obj{ewai-} \qu{to sit down} \\
  \obj{əturu-} \qu{to talk} &  \obj{epɨ-} \qu{to bathe (\gl{intr})} \\
 \obj{əməmɨ-} \qu{to enter} & \obj{ehpa-} \qu{to bathe (\gl{intr})} \\
\obj{əənɨkɨ-} \qu{to sleep} &                                       \\
\mybottomrule
\end{tabular}
\end{table}