### 2023-08-23 documents/floats/go.tex

\begin{table}
\centering
\caption[Reflexes of \rc{ɨtə[mə]} \qu{to go}]{Reflexes of \rc{ɨtə[mə]} \qu{to go} \parencites[291]{cruz2005fonologia}[292]{triomeira1999}[195]{wayanatavares2005}[87]{gildea1994akuriyo}[80, 153]{alves2017arara}[27, 248]{hixkaryanaderby1985}[45, 62]{waiwaihawkins1998}[54, 80]{ikpengpacheco2001}[112, 374]{von1892bakairi}[181, 216]{maquiritaricaceres2011}[112]{meira1998proto}[5]{guerrero2016karihona}[168]{hoff1968carib}[139]{meira2006syntactic}[4]{caceres2018yawarana}[74]{muller1975mapoyo}[198]{mattei1994diccionario}[48, 50]{macushiabbott1991}[172]{garcia2006diccionario}[Spike Gildea]{pc}[6]{franchetto2002kuikuro}[99]{camargo2002lexico}[14]{lacerda1991waimiri}}
\label{tab:go}
\begin{tabular}[t]{@{}ll@{}}
\mytoprule
Language &             Form \\
\mymidrule
\kaxui    &     \obj{to[mo]} \\
\PWai     &    \rc{[ɨ]to[m]} \\
\hixka    &      \obj{[ɨ]to} \\
\waiwai   &   \obj{[e]to[m]} \\
\PPek     &      \rc{ɨtə[n]} \\
\bakairi  &      \obj{[ɨ]tə} \\
\arara    &         \obj{to} \\
\arara    &        \obj{ɨdo} \\
\ikpeng   &       \obj{aran} \\
\ikpeng   &        \obj{ero} \\
\carijo   &     \obj{tə[mə]} \\
\PTir     &   \rc{[ə]tə[mɨ]} \\
\trio     &      \obj{tə[n]} \\
\akuriyo  &  \obj{[ə]tə[mɨ]} \\
\wayana   &   \obj{[ɨ]tə[m]} \\
\apalai   &        \obj{ɨto} \\
\kalina   &       \obj{[ɨ]ʔ} \\
\kalina   &         \obj{to} \\
\maqui    &    \obj{ɨtə[mə]} \\
\ingariko &        \obj{ətə} \\
\pemon    &      \obj{[e]tə} \\
\macushi  &      \obj{[a]tɨ} \\
\panare   &      \obj{tə[n]} \\
\yawarana &         \obj{tə} \\
\mapoyo   &         \obj{tə} \\
\uxc      &      \obj{[e]te} \\
\yukpa    &         \obj{to} \\
\waimiri  &         \obj{ɨʔ} \\
\mybottomrule
\end{tabular}
\end{table}