### 2023-08-23 documents/floats/say.tex

\begin{table}
\centering
\caption[Reflexes of \rc{ka[ti]} \qu{to say}]{Reflexes of \rc{ka[ti]} \qu{to say} \parencites[267]{meira2005bakairi}[4]{meira2003bakairi}[48]{franchetto2008absolutivo}[209, 279]{ikpengpacheco2001}[80, 153]{alves2017arara}[182]{hixkaryanaderby1985}[113]{meira1998proto}[107]{koehn1986apalai}[26]{waiwaihawkins1998}[Spike Gildea]{pc}[66]{camargo2010wayana}[59]{macushiabbott1991}[123]{swiggers2010gramatica}[430]{courtz2008carib}[125]{akawaiocaesar2003}[102]{mattei1994diccionario}[63]{largo2011yukpa}}
\label{tab:say}
\begin{tabular}[t]{@{}ll@{}}
\mytoprule
Language &          Form \\
\mymidrule
\kaxui   &   \obj{ka[s]} \\
\PWai    &    \rc{ka[s]} \\
\hixka   &   \obj{ka[s]} \\
\waiwai  &   \obj{ka[s]} \\
\PPek    &       \rc{ge} \\
\PPek    &       \rc{ke} \\
\bakairi &      \obj{ge} \\
\bakairi &      \obj{ke} \\
\arara   &      \obj{ge} \\
\arara   &      \obj{ke} \\
\ikpeng  &      \obj{ge} \\
\ikpeng  &    \obj{k[e]} \\
\carijo  &      \obj{ka} \\
\PTir    &       \rc{ka} \\
\trio    &      \obj{ka} \\
\akuriyo &      \obj{ka} \\
\wayana  &   \obj{ka[i]} \\
\apalai  &  \obj{ka[ʃi]} \\
\kalina  &      \obj{ka} \\
\kapon   &      \obj{ka} \\
\pemon   &      \obj{ka} \\
\macushi &      \obj{ka} \\
\panare  &   \obj{ka[h]} \\
\uxc     &      \obj{ki} \\
\yukpa   &      \obj{ka} \\
\mybottomrule
\end{tabular}
\end{table}