### 2023-08-23 documents/config/cariban_language_names.tex
fix abbrevs & lg names

\usepackage{xspace}
\providecommand{\wbakairi}{Western Bakairi\xspace}
\providecommand{\ebakairi}{Eastern Bakairi\xspace}
\providecommand{\ingariko}{Ingarikó\xspace}
\providecommand{\akawaio}{Akawaio\xspace}
\providecommand{\patamona}{Patamona\xspace}
\providecommand{\yukpa}{Yukpa\xspace}
\providecommand{\japreria}{Japreria\xspace}
\providecommand{\carijo}{Karijona\xspace}
\providecommand{\mapoyo}{Mapoyo\xspace}
\providecommand{\yawarana}{Yawarana\xspace}
\providecommand{\panare}{Panare\xspace}
\providecommand{\maqui}{Ye'kwana\xspace}
\providecommand{\pemon}{Pemón\xspace}
\providecommand{\kapon}{Kapong\xspace}
\providecommand{\waimiri}{Waimiri-Atroari\xspace}
\providecommand{\macushi}{Macushi\xspace}
\providecommand{\waiwai}{Waiwai\xspace}
\providecommand{\hixka}{Hixkaryána\xspace}
\providecommand{\kaxui}{Werikyana\xspace}
\providecommand{\trio}{Tiriyó\xspace}
\providecommand{\akuriyo}{Akuriyó\xspace}
\providecommand{\apalai}{Apalaí\xspace}
\providecommand{\wayana}{Wayana\xspace}
\providecommand{\ikpeng}{Ikpeng\xspace}
\providecommand{\arara}{Arara\xspace}
\providecommand{\kalina}{Kari'ña\xspace}
\providecommand{\uxc}{Upper Xingu Carib\xspace}
\providecommand{\bakairi}{Bakairi\xspace}
\providecommand{\PC}{Proto-Cariban\xspace}
\providecommand{\PPar}{Proto-Parukotoan\xspace}
\providecommand{\PWai}{Proto-Waiwaian\xspace}
\providecommand{\PPek}{Proto-Pekodian\xspace}
\providecommand{\PXin}{Proto-Xinguan\xspace}
\providecommand{\PPP}{Proto-Pemong-Panare\xspace}
\providecommand{\PTar}{Proto-Taranoan\xspace}
\providecommand{\PTir}{Proto-Tiriyoan\xspace}