### 2023-08-23 documents/config/own_abbreviations.tex

\DeclareAcronym{1}{short=1,long=first~person,short-format=\scshape}
\DeclareAcronym{1+2}{short=1+2,long=first~person~inclusive,short-format=\scshape}
\DeclareAcronym{1+3}{short=1+3,long=first~person~exclusive,short-format=\scshape}
\DeclareAcronym{2}{short=2,long=second~person,short-format=\scshape}
\DeclareAcronym{3}{short=3,long=third~person,short-format=\scshape}
\DeclareAcronym{a}{short=A,long=agent-like~argument~of~canonical~transitive~verb,short-format=\scshape}
\DeclareAcronym{agt}{short=agt,long=agent,short-format=\scshape}
\DeclareAcronym{ana}{short=ana,long=anaphoric,short-format=\scshape}
\DeclareAcronym{attrz}{short=attrz,long=attributivizer,short-format=\scshape}
\DeclareAcronym{caus}{short=caus,long=causative,short-format=\scshape}
\DeclareAcronym{cert}{short=cert,long=certainty,short-format=\scshape}
\DeclareAcronym{com}{short=com,long=comitative,short-format=\scshape}
\DeclareAcronym{cont}{short=cont,long=continuative,short-format=\scshape}
\DeclareAcronym{cor}{short=cor,long=coreference,short-format=\scshape}
\DeclareAcronym{dem}{short=dem,long=demonstrative,short-format=\scshape}
\DeclareAcronym{des}{short=des,long=desiderative,short-format=\scshape}
\DeclareAcronym{detrz}{short=detrz,long=detransitivizer,short-format=\scshape}
\DeclareAcronym{dub}{short=dub,long=dubitative,short-format=\scshape}
\DeclareAcronym{evid}{short=evid,long=evidentiality,short-format=\scshape}
\DeclareAcronym{frust}{short=frust,long=frustrative,short-format=\scshape}
\DeclareAcronym{hsy}{short=hsy,long=hearsay/indirect~evidentiality,short-format=\scshape}
\DeclareAcronym{imm}{short=imm,long=immentiate~past,short-format=\scshape}
\DeclareAcronym{imp}{short=imp,long=imperative,short-format=\scshape}
\DeclareAcronym{inan}{short=inan,long=inanimate,short-format=\scshape}
\DeclareAcronym{inf}{short=inf,long=infinitive,short-format=\scshape}
\DeclareAcronym{intr}{short=intr,long=intermediate~past,short-format=\scshape}
\DeclareAcronym{ints}{short=ints,long=intensifier,short-format=\scshape}
\DeclareAcronym{lk}{short=lk,long=linker,short-format=\scshape}
\DeclareAcronym{loc}{short=loc,long=locative,short-format=\scshape}
\DeclareAcronym{med}{short=med,long=medial,short-format=\scshape}
\DeclareAcronym{neg}{short=neg,long=negation,short-format=\scshape}
\DeclareAcronym{nmlz}{short=nmlz,long=nominalizer,short-format=\scshape}
\DeclareAcronym{npst}{short=npst,long=non-past,short-format=\scshape}
\DeclareAcronym{obl}{short=obl,long=oblique,short-format=\scshape}
\DeclareAcronym{p}{short=p,long=patient-like~argument~of~canonical~transitive~verb,short-format=\scshape}
\DeclareAcronym{pfv}{short=pfv,long=perfective,short-format=\scshape}
\DeclareAcronym{pl}{short=pl,long=plural,short-format=\scshape}
\DeclareAcronym{pro}{short=pro,long=pronoun,short-format=\scshape}
\DeclareAcronym{prs}{short=prs,long=present,short-format=\scshape}
\DeclareAcronym{pst}{short=pst,long=past,short-format=\scshape}
\DeclareAcronym{rec}{short=rec,long=recent~past,short-format=\scshape}
\DeclareAcronym{rem}{short=rem,long=remote~past,short-format=\scshape}
\DeclareAcronym{s}{short=s,long=single~argument~of~canonical~intransitive~verb,short-format=\scshape}
\DeclareAcronym{sa}{short=S\textsubscript{A},long=S~marked~like~A,short-format=\scshape}
\DeclareAcronym{sp}{short=S\textsubscript{P},long=S~marked~like~P,short-format=\scshape}
\DeclareAcronym{sap}{short=sap,long=speech~act~participant,short-format=\scshape}
\DeclareAcronym{sup}{short=sup,long=supine,short-format=\scshape}
\DeclareAcronym{tam}{short=TAM,long=tense-aspect-mood,short-format=\scshape}
\DeclareAcronym{tr}{short=tr,long=transitive,short-format=\scshape}
\DeclareAcronym{uncert}{short=uncert,long=uncertainty,short-format=\scshape}
\DeclareAcronym{inter}{short=inter,long=interrogative,short-format=\scshape}
\DeclareAcronym{int}{short=int,long=intermediate~past,short-format=\scshape}
\DeclareAcronym{all}{short=all,long=allative,short-format=\scshape}
\DeclareAcronym{dat}{short=dat,long=dative,short-format=\scshape}