### 2023-08-23 documents/floats/pwaireg.tex

\begin{table}
\centering
\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai \parencites[150]{howard2001wrought}[189, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
\label{tab:pwaireg}
\begin{tabular}[t]{@{}lllllll@{}}
\mytoprule
{} &    \qu{to fall} &                  &                    &     \qu{to sleep} &                 &                    \\
{} &           \PWai &           \hixka &            \waiwai &             \PWai &          \hixka &            \waiwai \\
\mymidrule
\gl{1}   &  \rc{k-eɸurka-} &  \obj{k-ehurka-} &    \obj{k-eɸɨrka-} &   \rc{kɨ-wɨnɨkɨ-} &  \obj{kɨ-nɨkɨ-} &   \obj{kɨ-wɨnɨkɨ-} \\
\gl{2}   &  \rc{m-eɸurka-} &  \obj{m-ehurka-} &    \obj{m-eɸɨrka-} &    \rc{o-wɨnɨkɨ-} &  \obj{o-wnɨkɨ-} &   \obj{mɨ-wɨnɨkɨ-} \\
\gl{1+2} &  \rc{t-eɸurka-} &  \obj{t-ehurka-} &  \obj{t͡ʃ-eɸɨrka-} &  \rc{tɨt-wɨnɨkɨ-} &  \obj{tɨ-nɨkɨ-} &  \obj{tɨt-wɨnɨkɨ-} \\
\gl{3}   &  \rc{ɲ-eɸurka-} &  \obj{ɲ-ehurka-} &    \obj{ɲ-eɸɨrka-} &   \rc{nɨ-wɨnɨkɨ-} &  \obj{nɨ-nɨkɨ-} &   \obj{nɨ-wɨnɨkɨ-} \\
\mybottomrule
\end{tabular}
\end{table}