### 2023-11-24 documents/floats/ppekverbs.tex
xinguan forms

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[83, 111, 129, 170]{pinto2022bakairi}{meiraDBwbakairi}[76, 130, 137, 374, 397, 400]{von1892bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[Fael Ikpeng]{pc}}
\label{tab:ppekverbs}
\begin{tabular}[t]{@{}llllllll@{}}
\mytoprule
{} &                                         \PPek &             \PXin &          \arara &                       \ikpeng &                     \ebakairi &                     \wbakairi &                          \bakairi \\
\mymidrule
\qu{say}     &                                   \rc{wɨ-ge-} &       \rc{wɨ-ge-} &    \obj{wɨ-ge-} &                   \obj{ɨ-ge-} &                   \obj{u-ge-} &                   \obj{u-ge-} &                     \ort{u-ɣépa} \\
\qu{bathe}   &                                   \rc{w-ipɨ-} &       \rc{w-ibɨ-} &    \obj{w-ibɨ-} &     \obj{{\normalfont ∅}-ip-} &                    \obj{w-i-} &                    \obj{w-i-} &                                   \\
\qu{be-1}    &                                    \rc{w-ap-} &        \rc{w-ap-} &     \obj{w-ap-} &                             – &                \obj{w-a(wɨ)-} &                \obj{w-a(wə)-} &  \ort{töre-wakine}, \ort{wawí-le} \\
\qu{be-2}    &                                 \rc{w-et͡ʃi-} &     \rc{w-et͡ʃi-} &  \obj{w-it͡ʃi-} &  \obj{{\normalfont ∅}-et͡ʃi-} &                    \obj{w-i-} &                    \obj{w-i-} &                      \ort{wí-ta} \\
\qu{come}    &                              \rc{w-(əd-)epɨ-} &  \rc{w-(od-)ebɨ-} &    \obj{w-ebɨ-} &                 \obj{k-arep-} &                 \obj{k-əewɨ-} &                 \obj{k-əewə-} &                 \ort{kχ-aewí-le} \\
\qu{go down} &  \rc{w-ɨptə({\normalfont C}{\normalfont V})-} &     \rc{w-iptoŋ-} &  \obj{w-iptoŋ-} &  \obj{{\normalfont ∅}-iptoŋ-} &  \obj{{\normalfont ∅}-ɨtəgɨ-} &  \obj{{\normalfont ∅}-itəge-} &                \ort{kχ-itaké-he} \\
\qu{go}      &                                  \rc{w-ɨdən-} &      \rc{w-ɨdon-} &    \obj{w-ɨdo-} &                 \obj{k-aran-} &                   \obj{u-də-} &                   \obj{u-də-} &                       \ort{ū́ta} \\
\mybottomrule
\end{tabular}
\end{table}