### 2023-11-26 documents/floats/aku-1sa.tex
reworked extensions section

\begin{table}
\centering
\caption[Regular \akuriyo \gl{1}\gl{sa} markers]{Regular \akuriyo \gl{1}\gl{sa} markers \parencites[77, 79, 82, 84-87]{gildea1994akuriyo}{meiraDBakuriyo}}
\label{tab:aku-1sa}
\begin{tabular}[t]{@{}ll@{}}
\toprule
first person \obj{k-} & first person \obj{t͡ʃ-} \\
\midrule
\obj{əət͡ʃena-} ‘to cry’ & \obj{epɨ-} ‘to bathe (\gl{intr})’ \\
\obj{əməmɨ-} ‘to enter’ & \obj{ehpa-} ‘to bathe (\gl{intr})’ \\
\obj{əempa-} ‘to learn’ & \obj{etonema-} ‘to lie down’ \\
\obj{ətajiŋka-} ‘to run’ & \obj{ewai-} ‘to sit down’ \\
\obj{əənɨkɨ-} ‘to sleep’ & \obj{ekɨrɨka-} ‘to stay back’ \\
\obj{əturu-} ‘to talk’ & \obj{entapo-} ‘to yawn’ \\
\obj{əiwa-} ‘to tremble’ & \obj{etaka-} ‘to be satiated’ \\
\bottomrule
\end{tabular}
\end{table}