### 2023-11-26 documents/floats/bak-verbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \bakairi]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \bakairi \parencites{meiraDBebakairi}[83, 111-112, 170]{pinto2022bakairi}{meiraDBwbakairi}[130-131, 137, 374, 397, 400]{von1892bakairi}[4]{meira2003bakairi}}
\label{tab:bak-verbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \ebakairi & \wbakairi & \bakairi(19th century) \\
\midrule
\qu{say} & \obj{u-ge-} & \obj{u-ge-} & \ort{u-ɣépa} \\
\qu{bathe} & \obj{w-i-} & \obj{w-i-} &  \\
\qu{be-1} & \obj{w-a(wɨ)-} & \obj{w-a(wə)-} & \ort{wawí-le} \\
\qu{be-2} & \obj{w-i-} & \obj{w-i-} & \ort{wí-ta} \\
\qu{go down} & \obj{{\normalfont ∅}-ɨtəgɨ-} & \obj{{\normalfont ∅}-itəge-} & \ort{kχ-itaké-he} \\
\qu{go} & \obj{u-də-} & \obj{u-də-} & \ort{ū́ta} \\
\bottomrule
\end{tabular}
\end{table}