### 2023-11-26 documents/floats/hix-intro.tex

\begin{table}
\centering
\caption[Some \hixka verbs]{Some \hixka verbs \parencites[150, 510-511, 513, 520]{howard2001wrought}[198]{hixkaryanaderby1985}}
\label{tab:hix-intro}
\begin{tabular}[t]{@{}llllll@{}}
\toprule
 & ‘to fall’ & ‘to be afraid’ & ‘to walk’ & ‘to cut self’ & ‘to be’ \\
\midrule
\gl{1} & \obj{k-ehurka-} & \obj{k-oserʲehɨ-} & \obj{k-atarʲeknohɨ-} & \obj{k-atama-} & \obj{w-eʃe-} \\
\gl{2} & \obj{m-ehurka-} & \obj{m-oserʲehɨ-} & \obj{m-atarʲeknohɨ-} & \obj{m-atama-} & \obj{m-eʃe-} \\
\gl{1+2} & \obj{t-ehurka-} & \obj{t-oserʲehɨ-} & \obj{t-atarʲeknohɨ-} & \obj{t-atama-} & \obj{t-eʃe-} \\
\gl{3} & \obj{ɲ-ehurka-} & \obj{n-oserʲehɨ-} & \obj{n-atarʲeknohɨ-} & \obj{n-atama-} & \obj{n-eʃe-} \\
\bottomrule
\end{tabular}
\end{table}