### 2023-11-26 documents/floats/ikp-a.tex

\begin{table}
\centering
\caption[Lowering of mid vowels in \ikpeng]{Lowering of mid vowels in \ikpeng \parencites[9, 96]{desouza2010arara}[41]{ikpengpacheco1997}[21, 82, 98, 267, 284]{ikpengpacheco2001}[38]{desouza1988arara}[28, 102]{pinto2022bakairi}[131, 172, 187]{alves2017arara}[180, 184]{meira2005southern}}
\label{tab:ikp-a}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \arara & \ikpeng & \ebakairi \\
\midrule
‘dog’ & \obj{wokori} & \obj{okari} / \obj{akari} &  \\
‘path’ & \obj{enma} / \obj{onma} / \obj{anma} & \obj{anma} & \obj{ə̃wã} \\
‘to bring’ & \obj{enebɨ} & \obj{anep} & \obj{ene} \\
‘tree’ & \obj{jej} & \obj{jaj} & \obj{se} \\
‘what’ & \obj{odɨ} & \obj{arɨ} & \obj{ədɨ} \\
\bottomrule
\end{tabular}
\end{table}