### 2023-11-26 documents/floats/ikp-w.tex

\begin{table}
\centering
\caption[Loss of \rc{w} in \ikpeng]{Loss of \rc{w} in \ikpeng \parencites[143]{alves2017arara}[21, 164]{ikpengpacheco2001}[9]{desouza2010arara}[41]{ikpengpacheco1997}[44, 70]{souza1993arara}[40]{campetela1997analise}[118]{chagas2013verbo}}
\label{tab:ikp-w}
\begin{tabular}[t]{@{}lll@{}}
\toprule
 & \arara & \ikpeng \\
\midrule
‘DAT’ & \obj{wɨna} & \obj{ɨna} \\
‘dog’ & \obj{wokori} & \obj{okari} / \obj{akari} \\
‘capuchin monkey’ & \obj{tawe} & \obj{tae} \\
‘to sleep’ & \obj{wɨnkɨ} & \obj{ɨnkɨ} \\
‘to defecate’ & \obj{watke} & \obj{atke} \\
\bottomrule
\end{tabular}
\end{table}