### 2023-11-26 documents/floats/pek-reg.tex

\begin{table}
\centering
\caption[Regular Pekodian \gl{sa} verbs]{Regular Pekodian \gl{sa} verbs \parencites[4]{meira2003bakairi}[150]{alves2017arara}[52]{ikpengpacheco2001}}
\label{tab:pek-reg}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \ebakairi ‘to go up’ & \arara ‘to dance’ & \ikpeng ‘to run’ \\
\midrule
\gl{1} & \obj{k-əku-} & \obj{k-origu-} & \obj{k-aranme-} \\
\gl{2} & \obj{m-əku-} & \obj{m-origu-} & \obj{m-aranme-} \\
\gl{1+2} & \obj{kɨd-əku-} & \obj{kud-origu-} & \obj{kw-aranme-} \\
\gl{3} & \obj{n-əku-} & \obj{{\normalfont ∅}-origu} & \obj{{\normalfont ∅}-aranme-} \\
\bottomrule
\end{tabular}
\end{table}