### 2023-11-26 documents/floats/pek-verbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PPek \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}{meiraDBebakairi}[111-112, 129, 170]{pinto2022bakairi}[68]{ikpengpacheco1997}[4]{meira2003bakairi}[Fael Ikpeng]{pc}}
\label{tab:pek-verbs}
\begin{tabular}[t]{@{}lllll@{}}
\toprule
 & \PPek & \arara & \ikpeng & \ebakairi \\
\midrule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} & \obj{u-ge-} \\
\qu{bathe} & \rc{w-ipɨ-} & \obj{w-ibɨ-} & \obj{{\normalfont ∅}-ip-} & \obj{w-i-} \\
\qu{be-1} & \rc{w-apɨ-} & \obj{w-ap-} & – & \obj{w-a(wɨ)-} \\
\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{{\normalfont ∅}-et͡ʃi-} & \obj{w-i-} \\
\qu{come} & \rc{w-epɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} & \obj{k-əewɨ-} \\
\qu{go down} & \rc{w-ɨptə({\normalfont C}{\normalfont V})-} & \obj{w-iptoŋ-} & \obj{{\normalfont ∅}-iptoŋ-} & \obj{{\normalfont ∅}-ɨtəgɨ-} \\
\qu{go} & \rc{w-ɨdən-} & \obj{w-ɨdo-} & \obj{k-aran-} & \obj{u-də-} \\
\bottomrule
\end{tabular}
\end{table}