### 2023-11-26 documents/floats/ptir-reg.tex

\begin{table}
\centering
\caption[Regular \PTir \gl{sa} verbs]{Regular \PTir \gl{sa} verbs \parencites[292, 294]{triomeira1999}[87]{gildea1994akuriyo}}
\label{tab:ptir-reg}
\begin{tabular}[t]{@{}lllllll@{}}
\toprule
 & \multicolumn{3}{l}{‘to bathe (\gl{intr})’} & \multicolumn{3}{l}{‘to sleep’} \\
 & \PTir & \trio & \akuriyo & \PTir & \trio & \akuriyo \\
\midrule
\gl{1} & \rc{t͡ʃ-epɨ-} & \obj{s-epɨ-} & \obj{t͡ʃ-epɨ-} / \obj{t͡ʃ-ehpa-} & \rc{t-əənɨkɨ-} & \obj{t-əənɨkɨ-} & \obj{k-əənɨkɨ-} \\
\gl{2} & \rc{m-epɨ-} & \obj{m-epɨ-} & \obj{m-epɨ-} & \rc{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} & \obj{m-əənɨkɨ-} \\
\gl{1+2} & \rc{ke-epɨ-} & \obj{ke-epɨ-} & \obj{ke-epɨ-} & \rc{kɨt-əənɨkɨ-} & \obj{kɨt-əənɨkɨ-} & \obj{kəʔ-əənɨkɨ-} \\
\gl{3} & \rc{n-epɨ-} & \obj{n-epɨ-} & \obj{n-epɨ-} & \rc{n-əənɨkɨ-} & \obj{n-əənɨkɨ-} & \obj{n-əənɨkɨ-} \\
\bottomrule
\end{tabular}
\end{table}