### 2023-11-26 documents/floats/ptir-verbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PTir \parencites[292, 294, 339]{triomeira1999}[112-115]{meira1998proto}{meiraDBakuriyo}}
\label{tab:ptir-verbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \PTir & \trio & \akuriyo \\
\midrule
\qu{go} & \rc{w-ɨtəmɨ-} & \obj{wɨ-tən-} & \obj{ə-təmɨ-} \\
\qu{say} & \rc{wɨ-ka-} & \obj{wɨ-ka-} & \obj{wɨ-ka-} \\
\qu{come} & \rc{w-əʔepɨ-} & \obj{w-əepɨ-} & \obj{{\normalfont ∅}-eepɨ-} \\
\qu{be-1} & \rc{w-a-} & \obj{w-a-} & \obj{{\normalfont ∅}-a-} \\
\qu{be-2} & \rc{w-eʔi-} & \obj{w-ei-} & \obj{{\normalfont ∅}-eʔi-} \\
\bottomrule
\end{tabular}
\end{table}