### 2023-11-26 documents/floats/pwai-reg.tex

\begin{table}
\centering
\caption[Regular \PWai verbs]{Regular \qu{to fall} (\gl{sa}) and \qu{to sleep} (\gl{sp}) in \PWai \parencites[150]{howard2001wrought}[189-190, 196]{hixkaryanaderby1985}[30]{waiwaihawkins1998}[209, 211]{hawkins1953waiwai}}
\label{tab:pwai-reg}
\begin{tabular}[t]{@{}lllllll@{}}
\toprule
 & \multicolumn{3}{l}{‘to fall’} & \multicolumn{3}{l}{‘to sleep’} \\
 & \PWai & \hixka & \waiwai & \PWai & \hixka & \waiwai \\
\midrule
\gl{1} & \rc{k-eɸurka-} & \obj{k-ehurka-} & \obj{k-eɸɨrka-} & \rc{kɨ-wɨnɨkɨ-} & \obj{kɨ-nɨkɨ-} & \obj{kɨ-wɨnɨkɨ-} \\
\gl{2} & \rc{m-eɸurka-} & \obj{m-ehurka-} & \obj{m-eɸɨrka-} & \rc{o-wɨnɨkɨ-} & \obj{o-wnɨkɨ-} & \obj{mɨ-wɨnɨkɨ-} \\
\gl{1+2} & \rc{t-eɸurka-} & \obj{t-ehurka-} & \obj{t͡ʃ-eɸɨrka-} & \rc{tɨt-wɨnɨkɨ-} & \obj{tɨ-nɨkɨ-} & \obj{tɨt-wɨnɨkɨ-} \\
\gl{3} & \rc{ɲ-eɸurka-} & \obj{ɲ-ehurka-} & \obj{ɲ-eɸɨrka-} & \rc{nɨ-wɨnɨkɨ-} & \obj{nɨ-nɨkɨ-} & \obj{nɨ-wɨnɨkɨ-} \\
\bottomrule
\end{tabular}
\end{table}