### 2023-11-26 documents/floats/tri-intro.tex

\begin{table}
\centering
\caption[Some \trio verbs]{Some \trio verbs \parencites[292, 294]{triomeira1999}[274]{triocarlin2004}}
\label{tab:tri-intro}
\begin{tabular}[t]{@{}llllll@{}}
\toprule
 & ‘to sleep’ & ‘to see self’ & ‘to bathe (\gl{intr})’ & ‘to yawn’ & ‘to go’ \\
\midrule
\gl{1} & \obj{t-əənɨkɨ-} & \obj{t-əene-} & \obj{s-epɨ-} & \obj{s-entapo-} & \obj{wɨ-tən-} \\
\gl{2} & \obj{m-əənɨkɨ-} & \obj{m-əene-} & \obj{m-epɨ-} & \obj{m-entapo-} & \obj{mɨ-tən-} \\
\gl{1+2} & \obj{kɨt-əənɨkɨ-} & \obj{k-əene-} & \obj{ke-epɨ-} & \obj{k-entapo-} & \obj{kɨ-tən-} \\
\gl{3} & \obj{n-əənɨkɨ-} & \obj{n-əene-} & \obj{n-epɨ-} & \obj{n-entapo-} & \obj{nɨ-tən-} \\
\bottomrule
\end{tabular}
\end{table}