### 2023-11-26 documents/floats/xin-verbs.tex

\begin{table}
\centering
\caption[Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PXin]{Verbs preserving \gl{1}\gl{sa} \rc{w-} in \PXin \parencites[153, 200]{alves2017arara}[42, 80, 139, 209]{ikpengpacheco2001}[68]{ikpengpacheco1997}[Fael Ikpeng]{pc}}
\label{tab:xin-verbs}
\begin{tabular}[t]{@{}llll@{}}
\toprule
 & \PXin & \arara & \ikpeng \\
\midrule
\qu{say} & \rc{wɨ-ge-} & \obj{wɨ-ge-} & \obj{ɨ-ge-} \\
\qu{bathe} & \rc{w-ibɨ-} & \obj{w-ibɨ-} & \obj{{\normalfont ∅}-ip-} \\
\qu{be-1} & \rc{w-ap-} & \obj{w-ap-} & – \\
\qu{be-2} & \rc{w-et͡ʃi-} & \obj{w-it͡ʃi-} & \obj{{\normalfont ∅}-et͡ʃi-} \\
\qu{come} & \rc{w-(od)ebɨ-} & \obj{w-ebɨ-} & \obj{k-arep-} \\
\qu{go down} & \rc{w-iptoŋ-} & \obj{w-iptoŋ-} & \obj{{\normalfont ∅}-iptoŋ-} \\
\qu{go} & \rc{w-ɨdon-} & \obj{w-ɨdo-} & \obj{k-aran-} \\
\bottomrule
\end{tabular}
\end{table}